---
title: "Cleaning Batting Projections 2024"
output: html_document
date: "2023-02-06"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# bzip <- read.csv("bat_zips.csv")
bsteamer <- read.csv("~/Documents/GitHub/fantasy_baseball_draft_bot/R-stuff/2024/data/batc_2024.csv")
# bdc <- read.csv("bat_dc.csv")
# bbat <- read.csv("bat_bat.csv")
# batc <- read.csv("bat_atc_23.csv")
```

# Exploratory functions

```{r}
# nrow(bzip)
nrow(bsteamer)
# nrow(bdc)
# nrow(bbat)
# nrow(batc)
```

# Reorder largest to smallest

```{r}
# ordered <- c("bsteamer", "bzip", "batc", "bbat", "bdc")
```

The formula for Runs Created is as follows (roughly):

=((H+BB+HBP-CS)*(TB)+.26*(BB+HBP))+.52(SB))/(AB+BB+HBP))

Note

TB = SLG * AB

# Make Runs Created

## First, make TB

```{r}
bsteamer$TB <- with(bsteamer, SLG * AB)
# bzip$TB <- with(bzip, SLG * AB)
# batc$TB <- with(batc, SLG * AB)
# bbat$TB <- with(bbat, SLG * AB)
# bdc$TB <- with(bdc, SLG * AB)
```

## Next try making RC

```{r}
# bsteamer$RC <- with(bsteamer, ((H+BB+HBP-CS)*(TB+.26*(BB+HBP))+.52*(SB))/(AB+BB+HBP))
# bzip$RC <- with(bzip, ((H+BB+HBP-CS)*(TB+.26*(BB+HBP))+.52*(SB))/(AB+BB+HBP))

# If they have HBP and CS in the projections:
bsteamer$RC <- with(bsteamer, ((H+BB+HBP-CS)*(TB+.26*(BB))+.52*(SB))/(AB+BB))

# batc$RC <- with(batc, ((H+BB+HBP-CS)*(TB+.26*(BB))+.52*(SB))/(AB+BB))
# Without HBP and CS:
#batc$RC <- with(batc, ((H+BB)*(TB+.26*(BB))+.52*(SB))/(AB+BB))

# bbat$RC <- with(bbat, ((H+BB-CS)*(TB+.26*(BB))+.52*(SB))/(AB+BB))
# bdc$RC <- with(bdc, ((H+BB+HBP-CS)*(TB+.26*(BB+HBP))+.52*(SB))/(AB+BB+HBP))
```

Note, bbat and batc don't have HBP projected

# Weight the ratio stats

```{r}
bsteamer$wAVG <- bsteamer$AVG * bsteamer$AB
bsteamer$wOPS <- bsteamer$OPS * bsteamer$AB
```

## 2023

```{r}
batting <- bsteamer %>%
  select(Name, Team, PlayerId, AB, R, HR, RBI, RC, wAVG, wOPS, AVG, OPS)

names(batting)
```

## Write in the z-score function

```{r}
z_score <- function(x) {
  x <- as.numeric(x)
  m <- mean(x, na.rm = TRUE)
  s <- sd(x, na.rm = TRUE)
  z <- (x - m) / s
  z
}
```

## Use the z-score function

```{r}
batting_2 <- batting
non_cats_names <- c('Name', 'Team', 'PlayerId', 'AB', 'AVG', 'OPS')
non_cats <- which(names(batting_2) == non_cats_names)
temp <- apply(batting_2[, -non_cats], 2, z_score)
batting_2 <- cbind(batting_2[, non_cats], temp)
sums <- apply(batting_2[, c(7:12)], 1, sum)
batting_3 <- batting_2
batting_3$pr <- sums
batting_ordered <- batting_3[order(batting_3$pr, decreasing = TRUE), ]
batting_ordered$rank <- 1:nrow(batting_ordered)

# Repeat the calculations on only the top 300
# Normalize to this number to compare against batters better
temp <- left_join(batting_ordered[1:300, 1:3], batting, by = c("Name", 'Team', 'PlayerId'))
batting_2 <- temp
non_cats_names <- c('Name', 'Team', 'PlayerId', 'AB', 'AVG', 'OPS')
non_cats <- which(names(batting_2) == non_cats_names)
temp <- apply(batting_2[, -non_cats], 2, z_score)
batting_2 <- cbind(batting_2[, non_cats], temp)
sums <- apply(batting_2[, c(7:12)], 1, sum)
batting_3 <- batting_2
batting_3$pr <- sums
batting_ordered <- batting_3[order(batting_3$pr, decreasing = TRUE), ]
batting_ordered$rank <- 1:nrow(batting_ordered)

# add the raw data for context
batting_raw = batting[, -c(4, 9, 10)]
names(batting_raw)[-c(1:3)] <- paste(names(batting_raw)[-c(1:3)], '_raw', sep = '')
almost_done = left_join(batting_ordered, batting_raw, by = c("Name", 'Team', 'PlayerId'))
final_bat <- almost_done[, c(1:4, 15:20, 5:14)]
```

## See if it works

```{r}
batting_in_order <- batting_3[with(batting_3, order(batting_3$pr, decreasing = TRUE)), ]
head(batting_in_order)
```

```{r}
write.csv(batting_in_order, "~/Documents/GitHub/fantasy_baseball_draft_bot/R-stuff/2024/data/patc_pr.csv")
```

## Combining with pitching

```{r}
combined <- rbind(batting_ordered[, c(1:3, ncol(batting_ordered)-1)], pitching_ordered[, c(1:3, ncol(pitching_ordered)-2)])
shohei_combined <- sum(combined[combined$Name == 'Shohei Ohtani', 'pr'])
test <- combined %>% 
  distinct(Name, .keep_all = T)
test[test$Name == 'Shohei Ohtani', 'pr'] <- shohei_combined

joined <- left_join(test, id_map, by = c('PlayerId' = 'IDFANGRAPHS'))
final <- joined %>% 
  select(Name, Team, PlayerId, pr, POS) %>% 
  distinct(Name, .keep_all = T)
final <- final %>% 
  mutate(pos_2 = if_else(POS == '1B', '1B/3B',
                         if_else(POS == '3B', '1B/3B',
                                 if_else(POS == '2B', '2B/SS',
                                         if_else(POS == 'SS', '2B/SS',
                                                 if_else(POS == 'DH', 'UTIL',
                                                                 if_else(POS == 'SP', 'P',
                                                                         if_else(POS == 'RP', 'P',POS))))))))
final <- final %>% 
  select(Name, Team, pr, PlayerId)
names(final) <- c('name', 'team', 'points', 'IDFANGRAPHS')
# write.csv(final, 'players.csv')
```

# Join to ESPN

```{r}
espn <- read.csv("~/Documents/GitHub/fantasy_baseball_draft_bot/R-stuff/2024/data/espn_positions.csv")

adp_combined <- left_join(final, id_map[, c('IDFANGRAPHS', 'ESPNID')], by = 'IDFANGRAPHS')

pos <- left_join(adp_combined, espn[,c('PlayerId', 'position')], by = join_by('ESPNID' == 'PlayerId'))

pos[pos$name == 'Yoshinobu Yamamoto', 'position'] <- 'SP'
pos[pos$name == 'Jung Hoo Lee', 'position'] <- 'CF'
pos[pos$name == 'Evan Carter', 'position'] <- 'LF'
pos[pos$name == 'Colt Keith', 'position'] <- '3B-2B'
pos[pos$name == 'Wyatt Langford', 'position'] <- 'LF'
pos[pos$name == 'Jackson Holliday', 'position'] <- 'SS-2B'
pos[pos$name == 'Shota Imanaga', 'position'] <- 'SP'
pos[pos$name == 'Cristopher Sánchez', 'position'] <- 'SP'
pos[pos$name == 'Keaton Winn', 'position'] <- 'SP'
```

# Adjust PR by position, split position up if multi-eligible

```{r}
library(dplyr)
library(tidyr)

pos <- distinct(pos)
# Assuming 'df' is your initial data frame
df <- pos %>%
  separate_rows(position, sep = "-")

# Define n values for positions
n_values <- list(`RF` = 20, `DH` = 12, `1B` = 15, `LF` = 20, `3B` = 15,
                 `2B` = 15, `SS` = 15, `C` = 10, `SP` = 55, `RP` = 30)

# Define the function
get_nth_pr <- function(data, nth_values) {
  # Ensure nth_values is a named vector
  if (is.null(names(nth_values))) {
    stop("nth_values must be a named vector.")
  }
  
  # Initialize an empty data frame
  results_df <- data.frame(position = character(0), points = numeric(0), stringsAsFactors = FALSE)
  
  # Iterate through the positions and nth values
  for (pos in names(nth_values)) {
    nth_value <- nth_values[pos]
    nth_pr <- data %>%
      filter(position == pos) %>%
      arrange(desc(points)) %>%
      slice(nth_value) %>%
      pull(points)
    
    # If the position does not have enough data, return NA
    if (length(nth_pr) < 1) {
      nth_pr <- NA
    }
    
    # Fill the results data frame
    results_df <- rbind(results_df, data.frame(position = pos, replacement_pr = nth_pr))
  }
  
  return(results_df)
}
nth_values <- c("RF" = 20, "DH" = 12, "1B" = 15, "LF" = 20, "CF" = 20,
                "3B" = 15, "2B" = 15, "SS" = 15, "C" = 10, "SP" = 60,
                'RP' = 30)
result_matrix <- get_nth_pr(df, nth_values)

join_3 <- left_join(df, result_matrix[, c("position", 'replacement_pr')], by = "position")
join_3$adj_points <- join_3$points - join_3$replacement_pr

# join_3$name <- stri_trans_general(join_3$name, "Latin-ASCII")

# Get ready to join to adp data
join_4 <- left_join(join_3, id_map[, c('FANTPROSNAME', 'IDFANGRAPHS')], 
                    by = 'IDFANGRAPHS')

adp <- read.csv('/Users/williamfoote/Documents/GitHub/fantasy_baseball_draft_bot/R-stuff/2024/data/fantpros_adp_2024.csv')

join_5 <- left_join(join_4, adp[, c('Player', 'ESPN')], join_by('FANTPROSNAME' == 'Player'))
```

# Unique players df
```{r}
library(dplyr)

# Assuming df is your DataFrame and adj_points is your column of interest
unique_df <- join_5 %>%
  group_by(name, team, IDFANGRAPHS) %>%
  arrange(desc(adj_points), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup()

write.csv(join_5, "~/Documents/GitHub/fantasy_baseball_draft_bot/R-stuff/2024/data/full_results_2024.csv")
write.csv(unique_df, "~/Documents/GitHub/fantasy_baseball_draft_bot/R-stuff/2024/data/unique_results_2024.csv")
```

# Graphs

```{r}
library(ggplot2)
library(dplyr)

# Assuming join_3 is your data frame and it has the columns 'points' and 'position'
# If 'pr' is the column to be used for density and not 'points', replace 'points' with 'pr' in the code below

ggplot(join_3, aes(x = points, color = position, fill = position)) + 
  geom_density(alpha = 0.5) + 
  labs(x = "Points", y = "Density") +
  theme_minimal() +
  ggtitle("Density Plot of PR by Position") +
  scale_color_brewer(palette = "Set1") +  # Set color palette
  scale_fill_brewer(palette = "Set1")    # Ensure fill color matches line color
```

# Adp

```{r}

```

