---
title: "Cleaning Batting Projections 2024"
output: html_document
date: "2023-02-06"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# bzip <- read.csv("bat_zips.csv")
bsteamer <- read.csv("~/Documents/GitHub/fantasy_baseball_draft_bot/R-stuff/2024/data/batc_2024.csv")
# bdc <- read.csv("bat_dc.csv")
# bbat <- read.csv("bat_bat.csv")
# batc <- read.csv("bat_atc_23.csv")
```

# Exploratory functions

```{r}
# nrow(bzip)
nrow(bsteamer)
# nrow(bdc)
# nrow(bbat)
# nrow(batc)
```

# Reorder largest to smallest

```{r}
# ordered <- c("bsteamer", "bzip", "batc", "bbat", "bdc")
```

The formula for Runs Created is as follows (roughly):

=((H+BB+HBP-CS)*(TB)+.26*(BB+HBP))+.52(SB))/(AB+BB+HBP))

Note

TB = SLG * AB

# Make Runs Created

## First, make TB

```{r}
bsteamer$TB <- with(bsteamer, SLG * AB)
# bzip$TB <- with(bzip, SLG * AB)
# batc$TB <- with(batc, SLG * AB)
# bbat$TB <- with(bbat, SLG * AB)
# bdc$TB <- with(bdc, SLG * AB)
```

## Next try making RC

```{r}
# bsteamer$RC <- with(bsteamer, ((H+BB+HBP-CS)*(TB+.26*(BB+HBP))+.52*(SB))/(AB+BB+HBP))
# bzip$RC <- with(bzip, ((H+BB+HBP-CS)*(TB+.26*(BB+HBP))+.52*(SB))/(AB+BB+HBP))

# If they have HBP and CS in the projections:
bsteamer$RC <- with(bsteamer, ((H+BB+HBP-CS)*(TB+.26*(BB))+.52*(SB))/(AB+BB))

# batc$RC <- with(batc, ((H+BB+HBP-CS)*(TB+.26*(BB))+.52*(SB))/(AB+BB))
# Without HBP and CS:
#batc$RC <- with(batc, ((H+BB)*(TB+.26*(BB))+.52*(SB))/(AB+BB))

# bbat$RC <- with(bbat, ((H+BB-CS)*(TB+.26*(BB))+.52*(SB))/(AB+BB))
# bdc$RC <- with(bdc, ((H+BB+HBP-CS)*(TB+.26*(BB+HBP))+.52*(SB))/(AB+BB+HBP))
```

Note, bbat and batc don't have HBP projected

# Weight the ratio stats

```{r}
bsteamer$wAVG <- bsteamer$AVG * bsteamer$AB
bsteamer$wOPS <- bsteamer$OPS * bsteamer$AB
```

## 2023

```{r}
batting <- bsteamer %>%
  select(Name, Team, PlayerId, AB, R, HR, RBI, RC, wAVG, wOPS, AVG, OPS)

names(batting)
```

## Write in the z-score function

```{r}
z_score <- function(x) {
  x <- as.numeric(x)
  m <- mean(x, na.rm = TRUE)
  s <- sd(x, na.rm = TRUE)
  z <- (x - m) / s
  z
}
```

## Use the z-score function

```{r}
batting_2 <- batting
non_cats_names <- c('Name', 'Team', 'PlayerId', 'AB', 'AVG', 'OPS')
non_cats <- which(names(batting_2) == non_cats_names)
temp <- apply(batting_2[, -non_cats], 2, z_score)
batting_2 <- cbind(batting_2[, non_cats], temp)
sums <- apply(batting_2[, c(7:12)], 1, sum)
batting_3 <- batting_2
batting_3$pr <- sums
batting_ordered <- batting_3[order(batting_3$pr, decreasing = TRUE), ]
batting_ordered$rank <- 1:nrow(batting_ordered)

# Repeat the calculations on only the top 300
# Normalize to this number to compare against batters better
temp <- left_join(batting_ordered[1:300, 1:3], batting, by = c("Name", 'Team', 'PlayerId'))
batting_2 <- temp
non_cats_names <- c('Name', 'Team', 'PlayerId', 'AB', 'AVG', 'OPS')
non_cats <- which(names(batting_2) == non_cats_names)
temp <- apply(batting_2[, -non_cats], 2, z_score)
batting_2 <- cbind(batting_2[, non_cats], temp)
sums <- apply(batting_2[, c(7:12)], 1, sum)
batting_3 <- batting_2
batting_3$pr <- sums
batting_ordered <- batting_3[order(batting_3$pr, decreasing = TRUE), ]
batting_ordered$rank <- 1:nrow(batting_ordered)

# add the raw data for context
batting_raw = batting[, -c(4, 9, 10)]
names(batting_raw)[-c(1:3)] <- paste(names(batting_raw)[-c(1:3)], '_raw', sep = '')
almost_done = left_join(batting_ordered, batting_raw, by = c("Name", 'Team', 'PlayerId'))
final_bat <- almost_done[, c(1:4, 15:20, 5:14)]
```

## See if it works

```{r}
batting_in_order <- batting_3[with(batting_3, order(batting_3$pr, decreasing = TRUE)), ]
head(batting_in_order)
```

```{r}
write.csv(batting_in_order, "./data/batc_pr.csv")
```

## Combining with pitching

```{r}
combined <- rbind(batting_ordered[, c(1:3, ncol(batting_ordered)-1)], pitching_ordered[, c(1:3, ncol(pitching_ordered)-2)])
shohei_combined <- sum(combined[combined$Name == 'Shohei Ohtani', 'pr'])
test <- combined %>% 
  distinct(Name, .keep_all = T)
test[test$Name == 'Shohei Ohtani', 'pr'] <- shohei_combined

joined <- left_join(test, id_map, by = c('PlayerId' = 'IDFANGRAPHS'))
final <- joined %>% 
  select(Name, Team, PlayerId, pr, POS) %>% 
  distinct(Name, .keep_all = T)
final <- final %>% 
  mutate(pos_2 = if_else(POS == '1B', '1B/3B',
                         if_else(POS == '3B', '1B/3B',
                                 if_else(POS == '2B', '2B/SS',
                                         if_else(POS == 'SS', '2B/SS',
                                                 if_else(POS == 'DH', 'UTIL',
                                                                 if_else(POS == 'SP', 'P',
                                                                         if_else(POS == 'RP', 'P',POS))))))))
final <- final %>% 
  select(Name, Team, POS, pr)
names(final) <- c('name', 'team', 'position', 'points')
write.csv(final, 'players.csv')
```

# Combined projections

```{r}
full_20 <- read.csv('combined_2020_projections.csv')
full_21 <- read.csv('combined_2021_projections.csv')
full_22 <- read.csv('combined_2022_projections.csv')
names(full_20) <- names(full_21)
names(full_22) <- names(full_21)

combined_proj = rbind(full_20, full_21, full_22)
combined_proj = left_join(combined_proj, id_map[, c(9, 6, 19, 20)], by = c('playerid' = 'IDFANGRAPHS')) %>% 
  distinct()
write.csv(combined_proj, 'combined_projections.csv')

full_23 <- read.csv('combined_2023_projections.csv')
names(full_23) <- names(full_21)
combined_proj_23 <- left_join(full_23, id_map[, c(9, 6, 19, 20)], by = c('playerid' = 'IDFANGRAPHS')) %>% 
  distinct()
write.csv(combined_proj_23, 'combined_projections_23.csv')

```

## 2023 Combined projections

```{r}
fpros_adp <- read.csv('fpros_espn_adp_2023.csv')
joined_1 <- left_join(fpros_adp[, c('Player', 'ESPN')], id_map[, c('IDFANGRAPHS', 'FANTPROSNAME', 'ESPNID', 'ESPNNAME')], 
                      by = c('Player' = 'FANTPROSNAME'))

test <- joined_1 %>% 
  distinct(Player, ESPN, .keep_all = T)

test[1, c('IDFANGRAPHS', 'ESPNID', 'ESPNNAME')] <- c(13510, 32801, 'Jose Ramirez')
# test[266, c('ESPNID', 'ESPNNAME')] <- c(4424286, 'Spencer Torkelson')
# test[322, c('ESPNID', 'ESPNNAME')] <- c(41325, 'Aaron Ashby')
# test[32, c('ESPNID', 'ESPNNAME')] <- c(41199, 'Shane McClanahan')
# test[69, c('ESPNID', 'ESPNNAME')] <- c(36380, 'Jordan Romano')

bj <- left_join(batting_ordered, test, by = c('PlayerId' = 'IDFANGRAPHS'))
pj <- left_join(pitching_ordered, test, by = c('PlayerId' = 'IDFANGRAPHS'))

write.csv(bj, 'batting_for_combined_projections_23.csv')
write.csv(pj, 'pitching_for_combined_projections_23.csv')
```

