---
title: "MLB Data Analysis - Daily Projection Development"
output: html_document
date: "2023-04-02"
---

```{r}
# install.packages('baseballr')
library(baseballr)
```

# Load historical data

## Game ID's

We need this to join to game log data by player.

```{r}
# Repeat this for all dates in time frame
game_data_apr_1 = mlb_game_pks(as.Date('2023-04-01'))
```

Joining game log data to player data will have to be something like:
  #1. Get the opposing team
    - if equal to the player's team, then it's a home game
    - else it's at the opposing team
    - join team abbreviation (NYY or @TOR) to park id somehow
  #2. Find the opposing starting pitcher?
    - maybe get probable pitcher function
    - join pitcher logs to game log and then join this to the batter logs
  #3. Get player game log to game id: 
    - Match location (using #1), date (assume only one game per day)
# Feature Engineering

## Park factors

```{r}
fg_park(2022)
# fg_park_hand(yr) gets the park factor for hitter handedness
```
# Get batter and pitcher game logs for each id that shows up in underdog

```{r}
batter_df = fg_batter_game_logs(15640, year = 2022)
pitcher_df = fg_pitcher_game_logs(playerid = 13125, year = 2022)
for (i in seq_len(nrow(final))) {
  if (is.na(final[i, 'slotName'][[1]]) | (final[i, 'proj_fpts'][[1]] < 500)) {
    1
  } else if (final[i, 'slotName'][[1]] == 'P') {
    this_pitcher_id = as.numeric(final[i, 'PlayerId'][[1]])
    for (j in c(2019, 2020, 2021, 2022)) {
      this_pitcher_df = fg_pitcher_game_logs(playerid = this_pitcher_id, year = j)
      pitcher_df = rbind(pitcher_df, this_pitcher_df, fill = TRUE)
    }
  } else {
    this_batter_id = as.numeric(final[i, 'PlayerId'][[1]])
    for (j in c(2019, 2020, 2021, 2022)) {
      this_batter_df = fg_batter_game_logs(this_batter_id, year = 2022)
      batter_df = rbind(batter_df, this_batter_df, fill = TRUE)
    }
  }
}

pitcher_df = pitcher_df %>% distinct()
batter_df = batter_df %>% distinct()
saveRDS(batter_df, '~/Documents/GitHub/fantasy_baseball_draft_bot/data/batter_game_logs.RDS')
saveRDS(pitcher_df, '~/Documents/GitHub/fantasy_baseball_draft_bot/data/pitcher_game_logs.RDS')

```

## Clean pitcher data

```{r}
pitcher_df <- readRDS('~/Documents/GitHub/fantasy_baseball_draft_bot/data/pitcher_game_logs.RDS')
pitcher_df <- pitcher_df %>% mutate(QS = ifelse((GS == 1) & (IP >= 6) & (ER <=3), 1, 0))
pitcher_df <- pitcher_df %>% mutate(underdog_fpts = W * 2 + QS * 3 + SO + (floor(IP)+((IP - floor(IP))*(1/3))) - ER)
```

## Clean batter data

```{r}
batter_df <- readRDS('~/Documents/GitHub/fantasy_baseball_draft_bot/data/batter_game_logs.RDS')
batter_df <- batter_df %>% mutate(underdog_fpts = 3 * `1B` + 6 * `2B` + 8 * `3B` + 10 * HR + 3 * BB + 3 * HBP + 2 * RBI + 2 * R + 4 * SB)
```

